---
title: "Visual QA — myRFunctions plots"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    fig_width: 8
    fig_height: 5
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  dpi = 120
)

# Knit from `dev/` (RStudio Knit) or from the package root:
# rmarkdown::render("dev/visual-qa.Rmd")
pkg_root <- if (file.exists("DESCRIPTION")) {
  normalizePath(".")
} else if (file.exists(file.path("..", "DESCRIPTION"))) {
  normalizePath("..")
} else {
  stop(
    "Could not find the package root. Knit this file from `dev/` ",
    "or render it from the package root.",
    call. = FALSE
  )
}

if (!requireNamespace("devtools", quietly = TRUE)) {
  stop("Install devtools, then knit again: install.packages(\"devtools\")", call. = FALSE)
}

devtools::load_all(pkg_root, quiet = TRUE)

library(dplyr)
library(tibble)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
```

Knit this file locally and look at the figures before pushing plot-related changes. It is **not** run by testthat or `R CMD check`.

It loads the **current working copy** via `devtools::load_all()`, so uncommitted edits are what you see.

**How to run**

- In RStudio: open `dev/visual-qa.Rmd` and click Knit.
- From the package root: `rmarkdown::render("dev/visual-qa.Rmd")`

The HTML is written next to this file and is gitignored.

# Simulated data

Two small datasets, both seeded so re-knits are comparable.

```{r data}
set.seed(42)

stim_group <- expand.grid(
  subj  = factor(1:12),
  Stim  = factor(c("Low", "Med", "High"), levels = c("Low", "Med", "High")),
  Group = factor(c("Control", "Treatment"))
) |>
  as_tibble() |>
  mutate(
    y = 10 +
      3 * (Stim == "Med") +
      6 * (Stim == "High") +
      4 * (Group == "Treatment") +
      2.5 * (Stim == "High" & Group == "Treatment") +
      rnorm(n(), sd = 2.5)
  )

mod_sg <- suppressWarnings(lmer(y ~ Stim * Group + (1 | subj), data = stim_group))
emm_sg <- emmeans(mod_sg, ~ Stim | Group)
pw_sg  <- pairs(emm_sg, adjust = "tukey")

set.seed(99)

stim_diet_sex <- expand.grid(
  subj = factor(1:10),
  Stim = factor(c("Low", "Med", "High"), levels = c("Low", "Med", "High")),
  Diet = factor(c("Chow", "HFD")),
  Sex  = factor(c("F", "M"))
) |>
  as_tibble() |>
  mutate(
    y = 8 +
      2.5 * (Stim == "Med") +
      5 * (Stim == "High") +
      3 * (Diet == "HFD") +
      1.5 * (Sex == "M") +
      2 * (Stim == "High" & Diet == "HFD") +
      rnorm(n(), sd = 2)
  )

mod_sds <- suppressWarnings(
  lmer(y ~ Stim * Diet * Sex + (1 | subj), data = stim_diet_sex)
)
emm_sds <- emmeans(mod_sds, ~ Stim | Diet * Sex)
pw_sds  <- pairs(emm_sds, adjust = "tukey")
```

# `bbmake_pairwise_plot()`

## Default (effect annotation on)

Look for: T-capped CIs, connecting lines, one effect-size label per panel at the top right, significance brackets, a zero line, and extra top padding so the annotation is not cramped.

```{r pairwise-default, fig.height=4.8}
bbmake_pairwise_plot(emm_sg, pw = pw_sg, model = mod_sg) +
  labs(
    title = "Default: annotate_effect = TRUE",
    x = "Stimulation",
    y = "Response"
  )
```

## Effect annotation off

Look for: no top-right effect label, and **less** empty space above the brackets than the plot above.

```{r pairwise-no-ann, fig.height=4.8}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  annotate_effect = FALSE
) +
  labs(
    title = "annotate_effect = FALSE",
    x = "Stimulation",
    y = "Response"
  )
```

## Custom `y_expand`

Look for: a larger gap above the brackets than the default annotated plot.

```{r pairwise-y-expand, fig.height=4.8}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  y_expand = c(0.02, 0.45)
) +
  labs(
    title = "y_expand = c(0.02, 0.45)",
    x = "Stimulation",
    y = "Response"
  )
```

## Annotation text size follows the theme

Look for: the top-right effect label scaling with `base_size`, along with axis text. The first two use complete themes; the third only changes geom text via `theme(geom)`.

```{r pairwise-theme-11, fig.height=4.8}
p_ann <- bbmake_pairwise_plot(emm_sg, pw = pw_sg, model = mod_sg)

p_ann +
  theme_grey(base_size = 11) +
  labs(
    title = "theme_grey(base_size = 11)",
    x = "Stimulation",
    y = "Response"
  )
```

```{r pairwise-theme-18, fig.height=4.8}
p_ann +
  theme_bw(base_size = 18) +
  labs(
    title = "theme_bw(base_size = 18)",
    x = "Stimulation",
    y = "Response"
  )
```

```{r pairwise-theme-geom, fig.height=4.8}
p_ann +
  theme(geom = element_geom(fontsize = 14)) +
  labs(
    title = "theme(geom = element_geom(fontsize = 14))",
    subtitle = "Axis text stays at the default; only geom text (the annotation) should grow",
    x = "Stimulation",
    y = "Response"
  )
```

`theme(text = element_text(size = ...))` does **not** change `geom_text()` size. That is ggplot2's split between theme text and geom text.

## Facet grid (two by-variables)

Look for: one panel per Diet × Sex cell, annotation and brackets in each panel.

```{r pairwise-facet-grid, fig.height=6.5, fig.width=9}
bbmake_pairwise_plot(emm_sds, pw = pw_sds, model = mod_sds) +
  labs(
    title = "Auto facet_grid(Diet ~ Sex)",
    x = "Stimulation",
    y = "Response"
  )
```

## Other layout switches

```{r pairwise-no-connect, fig.height=4.5}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  connect = FALSE
) +
  labs(title = "connect = FALSE", x = "Stimulation", y = "Response")
```

```{r pairwise-no-facets, fig.height=4.5}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  facets = FALSE
) +
  labs(
    title = "facets = FALSE (all groups overplotted)",
    x = "Stimulation",
    y = "Response"
  )
```

```{r pairwise-no-zero, fig.height=4.5}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  show_zero_line = FALSE
) +
  labs(title = "show_zero_line = FALSE", x = "Stimulation", y = "Response")
```

```{r pairwise-hide-ns, fig.height=4.5}
bbmake_pairwise_plot(
  emm_sg, pw = pw_sg, model = mod_sg,
  hide.ns = TRUE
) +
  labs(title = "hide.ns = TRUE", x = "Stimulation", y = "Response")
```

# `bb_add_pairwise()`

Add emmeans pairwise brackets to an existing ggplot. The motivating layout is a dodged bar plot: `Stim` is compared within each `Diet × Satiety` x-position, faceted by `Sex`.

```{r add-pairwise-data}
set.seed(42)

pr_motivation <- expand.grid(
  subj    = factor(1:8),
  Stim    = factor(c("Cont", "Opto")),
  Sex     = factor(c("Male", "Female")),
  Diet    = factor(c("Chow", "HFHS")),
  Satiety = factor(c("Fed", "Fasted"))
) |>
  as_tibble() |>
  mutate(
    Breakpoint = 8 +
      6 * (Stim == "Opto" & Sex == "Female" & Diet == "Chow" & Satiety == "Fed") +
      3 * (Stim == "Opto" & Sex == "Female" & Diet == "HFHS" & Satiety == "Fasted") +
      rnorm(n(), sd = 1.2)
  )

mod_pr <- lm(Breakpoint ~ Stim * Sex * Diet * Satiety, data = pr_motivation)
emm_pr <- emmeans(mod_pr, ~ Stim | Sex * Diet * Satiety)
pw_pr  <- pairs(emm_pr, reverse = TRUE)

col_width <- 0.8
```

## Dodged bars (fill comparisons)

Look for: a small bracket spanning the Cont vs Opto bars at each `Diet.Satiety` tick, not across x-axis categories; separate Female / Male panels; brackets sitting above the jittered points.

```{r add-pairwise-dodge, fig.height=5, fig.width=9}
pr_motivation |>
  ggplot(aes(
    x = interaction(Diet, Satiety),
    y = Breakpoint,
    fill = Stim
  )) +
  geom_bar(
    stat = "summary",
    fun = mean,
    width = col_width,
    position = position_dodge(width = col_width)
  ) +
  geom_point(
    position = position_jitterdodge(
      dodge.width = col_width,
      jitter.width = col_width / 3
    )
  ) +
  facet_wrap(~ Sex) +
  bb_add_pairwise(pw_pr) +
  labs(
    title = "bb_add_pairwise() on dodged bars",
    subtitle = "pairs(emmeans(~ Stim | Sex * Diet * Satiety)); hide.ns = FALSE",
    x = "Diet × Satiety",
    y = "Breakpoint"
  )
```

## Significant contrasts only

Look for: the same layout, but only the Female / Chow / Fed (and possibly Female / HFHS / Fasted) brackets remain.

```{r add-pairwise-hide-ns, fig.height=5, fig.width=9}
pr_motivation |>
  ggplot(aes(
    x = interaction(Diet, Satiety),
    y = Breakpoint,
    fill = Stim
  )) +
  geom_bar(
    stat = "summary",
    fun = mean,
    width = col_width,
    position = position_dodge(width = col_width)
  ) +
  geom_point(
    position = position_jitterdodge(
      dodge.width = col_width,
      jitter.width = col_width / 3
    )
  ) +
  facet_wrap(~ Sex) +
  bb_add_pairwise(pw_pr, hide.ns = TRUE) +
  labs(
    title = "bb_add_pairwise(hide.ns = TRUE)",
    x = "Diet × Satiety",
    y = "Breakpoint"
  )
```

# `bb_add_errorbar()`

Model-based intervals from `emmeans()` on an existing ggplot. Default `interval = "ci"` uses the emmeans confidence limits (backtransformed when the grid was created with `type = "response"`). `interval = "sem"` is estimate ± SE on the plotted scale.

## CI on dodged bars

Look for: T-capped error bars on each Cont/Opto bar, aligned with the dodge, plus the pairwise brackets from above.

```{r add-errorbar-dodge, fig.height=5, fig.width=9}
pr_motivation |>
  ggplot(aes(
    x = interaction(Diet, Satiety),
    y = Breakpoint,
    fill = Stim
  )) +
  geom_bar(
    stat = "summary",
    fun = mean,
    width = col_width,
    position = position_dodge(width = col_width)
  ) +
  geom_point(
    position = position_jitterdodge(
      dodge.width = col_width,
      jitter.width = col_width / 3
    )
  ) +
  facet_wrap(~ Sex) +
  bb_add_errorbar(emm_pr) +
  bb_add_pairwise(pw_pr, hide.ns = TRUE) +
  labs(
    title = "bb_add_errorbar() + bb_add_pairwise()",
    subtitle = "emmeans CIs (interval = \"ci\")",
    x = "Diet × Satiety",
    y = "Breakpoint"
  )
```

## Backtransformed CI vs SEM (Poisson / log link)

Look for: CI whiskers that are **asymmetric** around the mean (longer above than below on the count scale). SEM whiskers that are **symmetric**. The CI is what you want after `type = "response"`.

```{r add-errorbar-poisson, fig.height=4.8, fig.width=9}
set.seed(1)
pois_dat <- expand.grid(
  Stim  = factor(c("Cont", "Opto")),
  Group = factor(c("A", "B")),
  rep   = 1:16
) |>
  as_tibble() |>
  mutate(y = rpois(n(), lambda = ifelse(Stim == "Opto", 8, 3)))

mod_pois <- glm(y ~ Stim * Group, data = pois_dat, family = poisson)
emm_pois <- emmeans(mod_pois, ~ Stim | Group, type = "response")

p_ci <- ggplot(pois_dat, aes(x = Group, y = y, fill = Stim)) +
  geom_bar(
    stat = "summary", fun = mean,
    width = 0.8, position = position_dodge(width = 0.8)
  ) +
  bb_add_errorbar(emm_pois, interval = "ci") +
  labs(title = "interval = \"ci\" (backtransformed)", y = "Count")

p_sem <- ggplot(pois_dat, aes(x = Group, y = y, fill = Stim)) +
  geom_bar(
    stat = "summary", fun = mean,
    width = 0.8, position = position_dodge(width = 0.8)
  ) +
  bb_add_errorbar(emm_pois, interval = "sem") +
  labs(title = "interval = \"sem\" (± 1 SE)", y = "Count")

p_ci
```

```{r add-errorbar-poisson-sem, fig.height=4.8, fig.width=9}
p_sem
```

# `bb_ggbarplot()`

```{r bar-data}
set.seed(1)
bar_dat <- expand.grid(
  group     = factor(c("A", "B")),
  condition = factor(c("Cont", "Stim")),
  rep       = 1:8
) |>
  transform(
    score = 2 +
      as.numeric(condition) +
      as.numeric(group) +
      rnorm(length(group), sd = 0.5)
  )
```

## Default

Look for: mean + CI error bars, dodged bars, jittered points aligned to bars, a y = 0 line, `theme_bw()` with large text, Set1 fills.

```{r bar-default, fig.height=4.5}
bb_ggbarplot(bar_dat, x = "group", y = "score", fill = "condition") +
  labs(title = "Default bb_ggbarplot()", y = "Score")
```

## Points and zero line off

```{r bar-bare, fig.height=4.5}
bb_ggbarplot(
  bar_dat,
  x = "group",
  y = "score",
  fill = "condition",
  add_points = FALSE,
  add_hline = FALSE
) +
  labs(title = "add_points = FALSE, add_hline = FALSE", y = "Score")
```

## With `pw` (via `bb_add_pairwise()`)

Look for: the default bar plot plus a bracket spanning the dodged Cont vs Stim bars in each `group`.

```{r bar-pairwise, fig.height=4.5}
mod_bar <- lm(score ~ group * condition, data = bar_dat)
pw_bar <- emmeans(mod_bar, ~ condition | group) |> pairs(reverse = TRUE)

bb_ggbarplot(
  bar_dat,
  x = "group",
  y = "score",
  fill = "condition",
  pw = pw_bar
) +
  labs(title = "bb_ggbarplot(..., pw = pairs(emmeans(~ condition | group)))", y = "Score")
```

# Tables (quick glance)

Not plots, but easy to scan here so a broken table does not sneak through with a plot change.

```{r tables}
bbmake_model_table(mod_sg)
bbmake_pairwise_table(pw_sg, model = mod_sg)
```
